\documentclass[aspectratio=169]{beamer}

% Theme and Color Setup
\usetheme{Madrid}
\usecolortheme{whale}
\useinnertheme{rectangles}
\useoutertheme{miniframes}

... % (The rest of the preamble remains unchanged as in the provided template)

% Document Start
\begin{document}

\frame{\titlepage}

\begin{frame}[fragile]
    \frametitle{Introduction to Trends in Big Data}
    \begin{block}{Overview}
        The field of Big Data is rapidly evolving, impacting industries by influencing decision-making, improving customer engagement, and driving innovation. Understanding these trends is crucial for businesses to stay competitive.
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Trends in Big Data - Part 1}
    \begin{enumerate}
        \item \textbf{Data Democratization}
            \begin{itemize}
                \item Description: Making data accessible to non-technical users through user-friendly interfaces and self-service analytics tools.
                \item Example: Companies like Tableau and Power BI enable users to create visualizations without extensive coding knowledge.
            \end{itemize}

        \item \textbf{AI and Machine Learning Integration}
            \begin{itemize}
                \item Description: Integrating AI and ML with Big Data analytics to enhance insights and automate processes.
                \item Example: Predictive analytics in retail tailors marketing strategies based on consumer behavior using algorithms.
            \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Trends in Big Data - Part 2}
    \begin{enumerate}
        \setcounter{enumi}{2}  % Continue numbering from previous frame
        \item \textbf{Edge Computing}
            \begin{itemize}
                \item Description: Processing data near the source to improve response times and reduce bandwidth costs.
                \item Example: Smart cities use edge computing for real-time traffic management.
            \end{itemize}

        \item \textbf{Data Privacy and Security}
            \begin{itemize}
                \item Description: Prioritizing data privacy to protect sensitive information amid increasing regulations.
                \item Example: Implementation of GDPR and HIPAA marks investments in securing personal data.
            \end{itemize}

        \item \textbf{Real-time Analytics}
            \begin{itemize}
                \item Description: Transitioning from batch processing to real-time analytics for timely decision-making.
                \item Example: Financial institutions use real-time fraud detection systems.
            \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Trends in Big Data - Part 3}
    \begin{enumerate}
        \setcounter{enumi}{5}  % Continue numbering from previous frame
        \item \textbf{Data Quality Over Quantity}
            \begin{itemize}
                \item Description: Emphasizing accuracy, relevance, and timeliness of data.
                \item Example: Companies using data cleansing tools to ensure reliable datasets.
            \end{itemize}
    \end{enumerate}
    
    \begin{block}{Key Takeaways}
        \begin{itemize}
            \item Cross-industry Impact: Trends affect various sectors including finance, healthcare, and retail.
            \item Adaptation is Essential: Organizations must adapt to harness data potential.
            \item Innovation and Development: Continuous tech advancements shape the future of Big Data analytics.
        \end{itemize}
    \end{block}
    
    \begin{block}{References for Further Reading}
        - \textit{"The Data Warehouse Toolkit"} by Ralph Kimball \\
        - \textit{"Big Data: A Revolution That Will Transform How We Live, Work, and Think"} by Viktor Mayer-Schönberger \& Kenneth Cukier
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Emerging Technologies}
    \begin{block}{Overview}
        Emerging technologies are revolutionizing big data processing, involving key advancements such as AI, machine learning, and cloud computing. These technologies optimize data management and decision-making.
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Components of Emerging Technologies}
    \begin{enumerate}
        \item \textbf{Artificial Intelligence (AI)}
        \begin{itemize}
            \item Definition: Systems performing tasks requiring human intelligence (e.g., visual perception).
            \item Role: Algorithms analyze vast datasets for insights.
            \item Example: AI in healthcare to predict diseases and recommend treatments.
        \end{itemize}
        
        \item \textbf{Machine Learning (ML)}
        \begin{itemize}
            \item Definition: Algorithms enabling computers to learn and make predictions.
            \item Role: Enhances predictive analytics based on past data.
            \item Example: Retail data analysis for inventory management and marketing optimization.
        \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Continued Key Components}
    \begin{enumerate}
        \setcounter{enumi}{2} % Continue numbering from 2
        \item \textbf{Deep Learning}
        \begin{itemize}
            \item Definition: Uses deep neural networks to analyze various data types.
            \item Role: Processes unstructured data for meaningful insights.
            \item Example: Facial recognition technology utilizing deep learning systems.
        \end{itemize}

        \item \textbf{Natural Language Processing (NLP)}
        \begin{itemize}
            \item Definition: AI branch focusing on humans-computer communication.
            \item Role: Analyzes large volumes of natural language data.
            \item Example: Sentiment analysis to gauge public opinion on social media.
        \end{itemize}

        \item \textbf{Automated Data Processing}
        \begin{itemize}
            \item Definition: Automates data collection, cleaning, and integration.
            \item Role: Reduces human error, increases efficiency.
            \item Example: ETL tools for simplified data movement into warehouses.
        \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Points and Conclusion}
    \begin{itemize}
        \item \textbf{Integration of AI/ML}: Enhances data-driven decision-making across industries.
        \item \textbf{Adaptability}: Organizations must continuously evolve with these technologies.
        \item \textbf{Business Impact}: Leveraging AI and ML improves efficiency, reduces costs, and enhances customer experiences.
    \end{itemize}

    \begin{block}{Conclusion}
        Emerging technologies like AI and machine learning are essential in transforming big data into actionable intelligence for future success.
    \end{block}

    \begin{block}{Diagram Suggestion}
        A flowchart depicting the relationship between big data, AI, ML, and their applications.
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Data Processing Methods}
    \begin{block}{Overview}
        Data processing methods have evolved significantly to meet the challenges of big data.
        This slide explores new methodologies, particularly focusing on automation and real-time analytics.
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Concepts: Automation in Data Processing}
    \begin{itemize}
        \item \textbf{Definition}: Automation involves using technology to perform tasks without human intervention.
        \item \textbf{Benefits}:
            \begin{itemize}
                \item Reduces manual errors
                \item Increases processing speed
                \item Enhances scalability
            \end{itemize}
        \item \textbf{Examples of Automation Tools}:
            \begin{itemize}
                \item \textbf{Apache Kafka}: A distributed streaming platform enabling real-time data processing.
                \item \textbf{Apache NiFi}: A tool designed to automate the flow of data between systems.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Concepts: Real-Time Analytics}
    \begin{itemize}
        \item \textbf{Definition}: Real-time analytics refers to analyzing data as it is created or received.
        \item \textbf{Benefits}:
            \begin{itemize}
                \item Immediate insights for better decision-making
                \item Ability to respond to events as they happen
            \end{itemize}
        \item \textbf{Examples}:
            \begin{itemize}
                \item \textbf{Streaming Data Analysis}: Analyzing live data streams from IoT devices.
                \item \textbf{Financial Services}: Real-time fraud detection systems flagging suspicious activities instantly.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Points & Conclusion}
    \begin{block}{Key Points to Emphasize}
        \begin{itemize}
            \item \textbf{Importance of Scalability}: Automated systems can efficiently process large datasets.
            \item \textbf{Integration with AI and Machine Learning}: Enhances automation capabilities by learning from data patterns.
        \end{itemize}
    \end{block}
    
    \begin{block}{Example Illustration}
        \begin{lstlisting}
[Data Source] --> [Ingesting Layer: Apache NiFi] --> 
[Processing Layer: Apache Kafka/Spark] --> 
[Analytics Layer: Real-time dashboards/AI Models]
        \end{lstlisting}
    \end{block}

    \begin{block}{Conclusion}
        The evolution of data processing methods is essential for leveraging big data, enhancing efficiency, accuracy, and insights.
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Case Studies in Big Data Technologies}
    \begin{itemize}
        \item Explore real-world applications of new technologies in big data
        \item Analyze successful implementations across various sectors
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Introduction to Case Studies}
    \begin{itemize}
        \item Focus on the integration of technologies in the big data ecosystem
        \item Highlight innovative applications and their business impacts
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Case Study 1: Retail & Inventory Management at Walmart}
    \begin{block}{Overview}
        Walmart employs big data analytics to manage its vast inventory across thousands of stores worldwide.
    \end{block}
    \begin{itemize}
        \item \textbf{Technologies Used:} Hadoop, real-time analytics tools
        \item \textbf{Process:} Analyze sales data and supply chain feedback for inventory optimization
    \end{itemize}
    \begin{block}{Results}
        \begin{itemize}
            \item Increased inventory turnover
            \item Enhanced customer satisfaction through product availability
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Case Study 2: Healthcare Data Analysis at Mount Sinai Health System}
    \begin{block}{Overview}
        Leveraging electronic health records (EHRs) to improve patient outcomes.
    \end{block}
    \begin{itemize}
        \item \textbf{Technologies Used:} Apache Spark, machine learning algorithms
        \item \textbf{Process:} Analyze patient data for treatment effectiveness and disease patterns
    \end{itemize}
    \begin{block}{Results}
        \begin{itemize}
            \item 15\% reduction in hospital readmissions
            \item Targeting high-risk patients for clinical trials
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Case Study 3: Financial Fraud Detection at PayPal}
    \begin{block}{Overview}
        Real-time fraud detection and prevention using big data analytics.
    \end{block}
    \begin{itemize}
        \item \textbf{Technologies Used:} Apache Kafka, machine learning models
        \item \textbf{Process:} Continuous monitoring of transactions for suspicious behavior
    \end{itemize}
    \begin{block}{Results}
        \begin{itemize}
            \item Significant reduction in fraudulent transactions
            \item Increased consumer trust in transaction security
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Points to Emphasize}
    \begin{itemize}
        \item Diverse applications across various sectors
        \item Enhanced operational efficiency and customer experiences
        \item Importance of emerging technologies like machine learning
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Conclusion}
    \begin{itemize}
        \item Case studies highlight the transformative potential of big data technologies
        \item Organizations can uncover valuable insights to drive strategic decisions
        \item Embracing these innovations fosters growth and competitiveness
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Impact on Industry - Overview}
    \begin{block}{Overview}
        Big Data is reshaping industries by enhancing decision-making, optimizing operations, and predicting future trends. The increasing volume, variety, and velocity of data generated offers unparalleled opportunities and challenges.
    \end{block}
    The impacts on finance, healthcare, and retail illustrate the transformative potential of Big Data.
\end{frame}

\begin{frame}[fragile]
    \frametitle{Impact on Industry - Finance}
    \begin{itemize}
        \item \textbf{Risk Management:} 
        Big Data analytics allows financial institutions to analyze vast datasets to better identify and mitigate risks. Predictive analytics can forecast market trends and potential risks, improving investment strategies.
        
        \item \textbf{Example:} 
        JPMorgan Chase utilizes machine learning algorithms to enhance fraud detection systems, analyzing transaction patterns in real-time to flag anomalies.

        \item \textbf{Personalized Banking:} 
        Banks use data analytics to tailor marketing efforts and service offerings to individual customer behaviors and preferences. 

        \item \textbf{Key Point:} 
        Data-driven insights can lead to increased customer satisfaction and loyalty.
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Impact on Industry - Healthcare and Retail}
    \begin{block}{Healthcare Industry}
        \begin{itemize}
            \item \textbf{Predictive Analytics:} 
            Healthcare providers leverage Big Data to predict patient outcomes and improve treatment protocols. Analyzing historical data helps identify risk factors for diseases.
            
            \item \textbf{Example:} 
            Hospitals employing predictive algorithms can reduce readmission rates by identifying patients at higher risk and providing tailored follow-up care.

            \item \textbf{Operational Efficiency:} 
            Big Data tools optimize hospital operations, such as scheduling and resource allocation, leading to cost reductions.

            \item \textbf{Key Point:} 
            Enhanced insights into patient care can lead to better clinical outcomes and operational efficiencies.
        \end{itemize}
    \end{block}

    \begin{block}{Retail Industry}
        \begin{itemize}
            \item \textbf{Customer Insights and Personalization:} 
            Retailers analyze shopping patterns, preferences, and feedback to customize marketing campaigns, reduce churn, and enhance customer experience.

            \item \textbf{Example:} 
            Amazon uses recommendation algorithms based on customer browsing and purchasing history to personalize product suggestions, significantly increasing sales.

            \item \textbf{Inventory Management:} 
            Big Data analytics provides real-time inventory tracking and demand forecasting, helping retailers reduce waste and ensure availability.

            \item \textbf{Key Point:} 
            The ability to predict consumer demand improves supply chain efficiency and profitability.
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Conclusion}
    Big Data is revolutionizing the way industries operate by providing valuable insights that lead to improved decision-making, enhanced customer experiences, and operational efficiencies. Organizations that harness these insights can achieve a competitive edge in their respective markets.

    \begin{itemize}
        \item \textbf{Statistics to Remember:} 
        As of recent estimates, 2.5 quintillion bytes of data are generated every day!
        
        \item \textbf{Interactivity:} 
        Encourage students to think of other industries that could benefit from Big Data analytics and discuss potential applications.
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Ethical Considerations - Overview}
    % Brief overview of the ethical implications in big data
    As big data technologies evolve across sectors, it is crucial to examine ethical considerations, particularly regarding data collection, storage, analysis, and sharing, amidst regulations like GDPR.
    
    \begin{itemize}
        \item Emphasis on user privacy and informed consent.
        \item Importance of data security against breaches.
        \item Necessity to address bias and fairness in algorithms.
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Ethical Issues in Big Data}
    % Overview of key ethical issues
    \begin{enumerate}
        \item \textbf{Privacy:} Concerns about individual rights regarding personal data collection. Are users informed?
        \item \textbf{Informed Consent:} Ensuring explicit consent is obtained for data processing.
        \item \textbf{Data Security:} Ethical obligation to protect sensitive data from breaches.
        \item \textbf{Bias and Fairness:} Risks from algorithms perpetuating existing biases, affecting recruitment and justice.
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{GDPR Compliance and Ethical Data Practices}
    % Discussing the GDPR principles and implications
    \begin{block}{GDPR Overview}
        \begin{itemize}
            \item \textbf{Right to Access:} Individuals can access their personal data and its usage.
            \item \textbf{Right to be Forgotten:} Option to delete data under specific conditions.
            \item \textbf{Data Minimization:} Collect only necessary data for its intended purpose.
        \end{itemize}
    \end{block}

    \begin{block}{Implementing GDPR Principles}
        Organizations should:
        \begin{itemize}
            \item Conduct regular audits for compliance.
            \item Provide clear privacy policies for data practices.
            \item Appoint Data Protection Officers for oversight.
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Examples of Ethical Dilemmas}
    % Highlighting real-world ethical dilemmas
    \begin{itemize}
        \item \textbf{Facial Recognition Technology:} Raises surveillance and consent issues often overlooked by users.
        \item \textbf{Targeted Advertising:} Can lead to manipulative practices based on behavior without clear consent.
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Final Thoughts}
    % Summary and future implications
    \begin{itemize}
        \item Ethical adherence fosters trust and enhances institutional reputation.
        \item The evolving big data landscape necessitates continuous evaluation of ethical standards.
        \item Organizations must proactively tackle ethical dilemmas to cultivate a data ethics culture.
    \end{itemize}
    
    As we advance in the realm of big data, prioritizing ethical considerations and compliance will be essential for responsible data practices.
\end{frame}

\begin{frame}[fragile]
    \frametitle{Future Directions in Big Data}
    \begin{block}{Introduction}
        As we look towards the future of big data, several promising advancements are anticipated to transform data processing and analytical techniques. Understanding these trends is crucial for professionals to remain relevant in an ever-evolving technological landscape.
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Trends and Advancements - Part 1}
    \begin{enumerate}
        \item \textbf{AI and Machine Learning Integration}
            \begin{itemize}
                \item Enhances predictive analytics and automates complex data interpretation.
                \item Example: Netflix uses ML models to recommend content based on viewing history.
            \end{itemize}
        
        \item \textbf{Enhanced Real-time Data Processing}
            \begin{itemize}
                \item Stream processing technologies enable data analysis as it is generated.
                \item Example: Financial institutions detect fraudulent transactions instantly.
            \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Trends and Advancements - Part 2}
    \begin{enumerate}
        \setcounter{enumi}{2} % Continue the numbering
        \item \textbf{Increased Use of Cloud Services}
            \begin{itemize}
                \item Cloud computing offers scalable storage and high-performance computing.
                \item Example: Businesses use AWS and Google Cloud for massive datasets.
            \end{itemize}
        
        \item \textbf{Privacy-Preserving Data Analysis}
            \begin{itemize}
                \item Techniques like federated learning maintain user privacy.
                \item Example: Google improves models without transferring sensitive data.
            \end{itemize}
        
        \item \textbf{Edge Computing}
            \begin{itemize}
                \item Processes data near the source to reduce latency and bandwidth.
                \item Example: Smart manufacturing analyzes sensor data to optimize production.
            \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Points and Conclusion}
    \begin{block}{Key Points to Emphasize}
        \begin{itemize}
            \item \textbf{Interconnectivity}: AI, real-time processing, and cloud services will lead to more interconnected data ecosystems.
            \item \textbf{Data Democratization}: A broader range of stakeholders will engage in data analysis.
            \item \textbf{Ethical Considerations}: Advancements must adhere to ethical guidelines, particularly about privacy.
        \end{itemize}
    \end{block}
    
    \begin{block}{Conclusion}
        The future of big data processing is dynamic, driven by advanced technologies that promise to enhance efficiency, insights, and ethical practices. Staying ahead of these trends is essential for professionals venturing into the data landscape.
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Preparing for Change}
    Strategies for professionals to stay relevant and adapt to the continually evolving landscape of big data.
\end{frame}

\begin{frame}[fragile]
    \frametitle{Overview}
    \begin{itemize}
        \item The landscape of big data is continually evolving due to:
        \begin{enumerate}
            \item Advancements in technology
            \item Changing consumer behavior
            \item Emerging regulatory frameworks
        \end{enumerate}
        \item Professionals must embrace change and develop relevant skills.
        \item This slide outlines practical strategies for adapting to changes in the field.
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Strategies for Professionals (1)}
    \begin{enumerate}
        \item \textbf{Continuous Learning}
        \begin{itemize}
            \item Engage in professional development through online courses and certifications.
            \item Resource Example: Platforms like Coursera and edX.
        \end{itemize}

        \item \textbf{Networking and Community Engagement}
        \begin{itemize}
            \item Join communities such as IEEE Big Data or ACM SIGKDD.
            \item Collaborate to gain insights and innovative solutions.
        \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Strategies for Professionals (2)}
    \begin{enumerate}[resume]
        \item \textbf{Embrace Emerging Technologies}
        \begin{itemize}
            \item Integrate AI and machine learning frameworks.
            \item Example Use Case: AI for predictive analytics.
        \end{itemize}

        \item \textbf{Adapt to Regulatory Changes}
        \begin{itemize}
            \item Understand regulations like GDPR, HIPAA, and CCPA.
            \item Regularly review compliance protocols.
        \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Strategies for Professionals (3)}
    \begin{enumerate}[resume]
        \item \textbf{Diversify Your Skill Set}
        \begin{itemize}
            \item Develop proficiency in programming languages (Python, R, SQL).
            \item Enhance soft skills for effective communication.
        \end{itemize}

        \item \textbf{Implement Agile Methodologies}
        \begin{itemize}
            \item Use agile practices for project flexibility.
            \item Example Methodology: Scrum framework for adaptive planning.
        \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Visual Example: Skill Development Roadmap}
    \begin{block}{Diagram Concept}
        A flowchart illustrating the progression from foundational knowledge to advanced expertise in big data tools, technologies, and compliance.
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Key Points to Emphasize}
    \begin{itemize}
        \item Continuous adaptation and learning are critical for success in big data.
        \item Networking enhances collaborative knowledge and innovative solutions.
        \item Keeping abreast of trends and regulations is essential to remain competitive.
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Conclusion - Overview}
    \begin{block}{Key Takeaways from Week 11}
        \begin{enumerate}
            \item Growing Importance of Big Data
            \item Emerging Technologies
            \item Data Privacy and Ethics
            \item Data Democratization
            \item Interdisciplinary Approach
            \item Focus on Real-Time Analytics
            \item Sustainability and Big Data
        \end{enumerate}
    \end{block}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Conclusion - Key Insights}
    \begin{itemize}
        \item Big data is becoming an essential driver of innovation across industries.
        \item Technologies like AI, ML, and edge computing are transforming data analytics.
        \item Data privacy regulations like GDPR are crucial for ethical data usage.
        \item Organizations are focusing on making data accessible to non-technical users (data democratization).
        \item An interdisciplinary approach fosters collaboration across fields to enhance big data insights.
        \item Real-time analytics are critical for immediate business decision-making.
        \item Sustainability is a growing concern, prompting companies to optimize energy usage in data management.
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Conclusion - Future Directions}
    \begin{block}{Summary}
        In summary, big data is set to shape our future, presenting both opportunities and challenges. Professionals need to stay updated on trends and adapt their strategies accordingly to remain relevant.
    \end{block}
    
    \begin{block}{Key Points to Remember}
        \begin{itemize}
            \item Innovation in big data is fueled by technology, ethics, and interdisciplinary collaboration.
            \item Adaptability and continuous learning are vital for professionals in the big data sphere.
        \end{itemize}
    \end{block}
    
    \begin{block}{Further Discussion}
        \begin{itemize}
            \item How can organizations implement these trends effectively?
            \item What are the potential risks associated with these advancements, and how can they be mitigated?
        \end{itemize}
    \end{block}
\end{frame}


\end{document}